---
title: "Final"
author: "Ngoc Nguyen, Craig Orman"
date: "4/29/2022"
output: html_document
---

```{r, knitr, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Selection
This is how we selected our data and ensured it would be small enough to go on github.
```{r, DataSelection, eval = FALSE}
HousingOriginal <- read.csv("properties_small_2017.csv")
smallData = HousingOriginal %>% select(yearbuilt, bathroomcnt, airconditioningtypeid, regionidcity, fireplacecnt,
                                       basementsqft, finishedsquarefeet15, landtaxvaluedollarcnt, typeconstructiontypeid, garagetotalsqft,
                                       latitude, longitude, taxvaluedollarcnt)
smallData = smallData %>% filter(bathroomcnt > 0, taxvaluedollarcnt > 999)
samllSampledData = sample_n(smallData, 1000000)
write.csv(samllSampledData,"properties_small_2017.csv", row.names = FALSE)
```

## Instantiation
```{r, Instantiation, warning=FALSE}
library(tidyverse)
library(ggmosaic)
library(maps)
library(ggplot2)
library(dplyr)
library(broom)
library(ggpubr)
HousingOriginal <- read.csv("properties_small_2017.csv")
```
## Proposal (Homework 11)

Data was acquired from a 2017 kaggle competition, [Zillow Prize: Zillow’s Home Value Prediction (Zestimate)](https://www.kaggle.com/competitions/zillow-prize-1). We specifically used the properties_2017.csv

Project is on [Github](https://github.com/ArgentCode/DS202-final-project).

Our plan is to use the taxvaluedollarcnt as the actual value of the house, and we plan on doing a variety of visualizations and assessments on both the explanatory variables and the response variable. We will be coming at this with a statistics approach as if we were preparing to make a model. We also hope to do a map to visualize the data and see what we can do for that style of visual with coloring and making interesting visuals.

We noticed the data is decently well cleaned. We suspect dealing with the large number of NA's will be a major challenge, but most of the cleaning has occured in the form of reducing. The original data was an almost 700 mb csv file we had to shrink down (Code above on methodology).

## Actual Report

Real estate and Housing prices are always in great flux. We have decided to use the 2017 Zillow competition data to try and predict prices for real estate using a variety of factors. The response variable we are using will be taxvaluedollarcnt. A continuous quantitative variable.

The response variables are:

Quantitative: 

yearbuilt, discrete

bathroomcnt, discrete with halves

fireplacecnt, discrete

basementsqft, continuous

finishedsquarefeet15, continuous

landtaxvaluedollarcnt, continuous

garagetotalsqft, continuous

Categorical:

airconditioningtypeid, nominal

regionidcity, nominal

typeconstructiontypeid, nominal

>__Slide 2__: 

### Ading expectation for variables:

Craig's expectations:

yearBuilt: Help explain taxvaluedollarcnt because newer houses will be nicer and have more modern ammenities like outlets

bathroomcnt:  Help explain taxvaluedollarcnt because more bathrooms usually is indicative of a bigger house

airconditioningtypeid: I have no idea what the distribution of airconditioning will be. But I am curious to find out.

regionIDCity: Help explain taxvaluedollarcnt because location is often one of the most important factors

fireplacecnt: I do not believe fireplaces are very common these days, so these will be more prevelant in older homes, making them less 
valuable.

------------------------------------

Ngoc's expectation:

basementsqft: I think basementsqft won't help explain taxvaluedollarcnt because a lot of houses do not have basements or has not finished it. I think larger houses has larger basement.

finishedsquarefeet15: Help explain taxvaluedollarcnt because if the house has large unfinished areas, the value is lesser than the one has larger 

landtaxvaluedollarcnt: the land tax usually base on the house's price. For example in Iowa, land tax is 1.53%. So the higher tax land, the higher 

typeconstructiontypeid:I don't think typeconstructiontypeid would well explain taxvaluedollarcnt because the type is not important but areas ( in square feet is). I would like to see if there is any time of construction has the price higher

garagetotalsqft:Bigger house has bigger garage so I think it make sense if this variable is well explain taxvaluedollarcnt. 

## Clean and visualize
First, how much data is invalid, and what is the structure of the data
```{r}
colSums(is.na(HousingOriginal))

str(HousingOriginal)
```

The variables airconditioningtypeid, fireplacecnt, basementsqft, finishedsquarefeet15, typeconstructiontypeid, garagetotalsqft all have a large majority of null values. fireplacecnt uses null to indicate that there is no fireplace. basementsqft uses null to indicate that some houses do not have basement (most of the houses in CA do not) and some basements haven't finished. However, garagetotalsqft and airconditioningtypeid have a category for no air conditioner so there is no obvious explanation for the lack of values. One possibility is those houses hasn't finished the garage and AC yet.  We still have ~300,000 observations so analysis should be able to continue.

All data is in numeric or integer format which is consistent with the Kaggle competition

Response Variable:

The response variable is our first investigation, we will be attempting to see if it is bell shaped and has a normal distribution. We see the data is quite spread out, with a standard deviation of $614,453 and a mean of $441,428. There are 103,994 observations of houses less than $100,000 and 443 observations of houses more than $10,000,000. So our data is incredibly spread out. 
```{r, ResponseVariable}
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt)) + geom_histogram()
HousingOriginal$taxvaluedollarcnt %>% summary() 
mean(HousingOriginal$taxvaluedollarcnt)
sd(HousingOriginal$taxvaluedollarcnt)
HousingOriginal[HousingOriginal$taxvaluedollarcnt > 10000000, ] %>% count()
HousingOriginal[HousingOriginal$taxvaluedollarcnt < 100000, ] %>% count()
```
>__Slide 4__:

# Visual summary 

### Quantitative Explanatory Variables:
```{r, yearbuilt}
HousingOriginal %>% ggplot(aes(x=yearbuilt)) + geom_boxplot()
HousingOriginal %>% 
  ggplot(aes(x=yearbuilt))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=yearbuilt)) + geom_point()
HousingOriginal$yearbuilt %>% summary()
```

2 sentences summary: Our median was 1963 with 75% of the data being between 1950 and 1981. Our data is skewwed left with a logical limit of 2017 when the Kaggle competition was taking place. There is a weak correlation between yearbuilt and taxvaluedollarcnt.

```{r}
HousingOriginal %>% ggplot(aes(x=bathroomcnt)) + geom_histogram()

HousingOriginal %>% 
  ggplot(aes(x=bathroomcnt))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=bathroomcnt)) + geom_point()
HousingOriginal$bathroomcnt %>% summary()
cor(HousingOriginal$taxvaluedollarcnt, HousingOriginal$bathroomcnt)
```

2 sentences summary: Data is unimodal and bell shaped with a right skew. There is a logical barrier at 0 as you cannot have negative bathrooms. Mean is right around 3 which is where its expected to be. There is a moderate-weak amount of correlation between bathroomcnt and taxvaluedollarcnt

```{r}
HousingOriginal %>% ggplot(aes(x=fireplacecnt)) + geom_histogram()

HousingOriginal %>% 
  ggplot(aes(x=fireplacecnt))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=fireplacecnt)) + geom_point()
HousingOriginal$fireplacecnt %>% summary()
```

2 sentences summary: As predicted, most of the observations have no fireplace. Results look logarithmic and very sporadic, there does not appear to be a correlation between fireplacecnt and taxvaluedollarcnt.

```{r}

HousingOriginal %>% ggplot(aes(x=basementsqft)) + geom_boxplot()

HousingOriginal %>% 
  ggplot(aes(x=basementsqft))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=basementsqft)) + geom_point()
HousingOriginal$basementsqft %>% summary()
cor(HousingOriginal$taxvaluedollarcnt, HousingOriginal$basementsqft)
```

2 sentences summary: The median square feet of a house's basement is 533 with 75% of the basement's area between 266 to 823.5 square feet. The graph skewed right. There does not appear to be a correlation between fireplacecnt and taxvaluedollarcnt.

```{r}

HousingOriginal %>% ggplot(aes(x=finishedsquarefeet15)) + geom_boxplot()

HousingOriginal %>% 
  ggplot(aes(x=finishedsquarefeet15))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=finishedsquarefeet15)) + geom_point()
HousingOriginal$finishedsquarefeet15 %>% summary()
```

2 sentences summary: The median of the data is at 2128 and 75% of the data is between 1681 and 2847. The data is highly concentrated on the left with outliers on the right. There are about 3-5 outliers of significance, and there does not appear to be any correlation between finishedsquarefeet15 and taxvaluedollarcnt.

```{r}
HousingOriginal %>% ggplot(aes(x=landtaxvaluedollarcnt)) + geom_boxplot()

HousingOriginal %>% 
  ggplot(aes(x=landtaxvaluedollarcnt))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=landtaxvaluedollarcnt)) + geom_point()
HousingOriginal$landtaxvaluedollarcnt %>% summary()
```

2 sentences summary: The median is 176659 with 75% of the data being between 80267 and 325091. The data is quite spread out with many outliers. There is a clear, almost dependent relationship between landtaxvaluedollarcnt and taxvaluedollarcnt

```{r}
HousingOriginal %>% ggplot(aes(x=garagetotalsqft)) + geom_boxplot()

HousingOriginal %>% 
  ggplot(aes(x=garagetotalsqft))+
  geom_line(stat="density", color="Red", size=1)+
  theme_bw()
HousingOriginal %>% ggplot(aes(x=taxvaluedollarcnt, y=garagetotalsqft)) + geom_point()
HousingOriginal$garagetotalsqft %>% summary()
cor(HousingOriginal$taxvaluedollarcnt, HousingOriginal$garagetotalsqft)
```

2 sentences summary: Median is 441 with 75% of the data being between 350 and 496. The data is very spread out, and there does not appear to be a correlation between garagetotalsqft and taxvaluedollarcnt.
 

### Categorical Explanatory Variables:
```{r}
HousingOriginal %>% ggplot(aes(x=airconditioningtypeid)) + geom_bar()
```

2 sentences summary: Almost all of the obsercations are of one type. Most of the other types aren't even visible.

```{r}
HousingOriginal %>% ggplot(aes(x=regionidcity)) + geom_histogram()
```

2 sentences summary: Data is surprisingly unimodal in a way. Large portion of the data is distributed across a few very similar codes. 

```{r}
HousingOriginal %>% ggplot(aes(x=typeconstructiontypeid)) + geom_bar()

``` 

2 sentences summary: Almost all of the observations are type 6. The rest of the variables are almost invisible.

Next, to get a general sense of what data we are talking about, we will map out the data on a plot. Zillow multiplied all longitudes and latitudes by 1*10e^6 so we accounted for this in the code.
```{r, MapData}
states <- map_data("state")
filter(states, region == "california") %>% ggplot(aes(x = long, y = lat)) +
  geom_path(aes(group = group)) +
  geom_point(data=HousingOriginal,aes(x=longitude/1000000, y=latitude/1000000, color="#E69F00")) +
  ggtitle("All data mapped against California")
```

>__Slide 6__: 
Expanding complexity

### Craig's Questions:
Variables: yearBuilt, bathroomcnt, airconditioningtypeid, regionIDCity, fireplacecnt

<!-- numeric vs numeric -->
1. What is the relationship between bathroomcnt and fireplacecnt

<!-- numeric vs categorical -->
2. What is the relationship between yearbuilt and airconditioningtypeid

<!-- numeric vs categorical -->
3. What is the relationship between firepalcecnt and yearbuilt

In modern architecture, fire places are not very common place, but do exist especially in nicer homes. In older architecture, they are more common for both stylistic and actual warmth purposes. Bathrooms are required for all homes to my knowledge, and having more of them is indicative of either a multi-family home or a nicer home. I would like to compare them and ensure these variables aren't dependent and that they are not just both indicative of similar properties.
```{r, BathFire}
CraigsHousing = HousingOriginal
#fireplacecnt had to have nas replaced with 0s for the plot
CraigsHousing$fireplacecnt = CraigsHousing$fireplacecnt %>% replace_na(0)
CraigsHousing[CraigsHousing$bathroomcnt > 10, ] %>% count()
CraigsHousing[CraigsHousing$bathroomcnt < CraigsHousing$fireplacecnt, ] %>% count()
CraigsHousing %>% ggplot(aes(x=bathroomcnt, y=fireplacecnt)) + geom_point()
cor(CraigsHousing$bathroomcnt, CraigsHousing$fireplacecnt)
```

With an r value of 0.087 we know these values are independent. We can also see from the data that fireplaces are pretty scattered across all bathrooms quantities. Also, there are 237 homes with more than 10 bathrooms, and 672 homes that have more fireplaces than bathrooms both of which I found interesting.

Air conditioning is a rather recent invention compared to some of these houses. With that in mind, I am quite curious what air conditioning systems are most common for different eras
```{r yearlyAirConditioning}
CraigsHousing %>% ggplot(aes(y=yearbuilt)) +
  geom_boxplot() +
  facet_grid(~airconditioningtypeid) +
  ggtitle("Yearbuilt by Airconditioningtypeid")
filter(CraigsHousing, airconditioningtypeid > 0) %>%
  ggplot(aes(y=taxvaluedollarcnt,
             x=yearbuilt,
             color=as.factor(airconditioningtypeid))) +
  geom_point()
```

1 Central, 3 Evaporative Cooler, 5 None, 9 Refrigeration, 11 Wall Unit, 12 Window Unit, 13 Yes
I have no idea what an evaporative cooler is but, they apparently existed very briefly. A lot more houses don't have air conditioning than I thought would. I also didn't expect central air in so many old houses. The comparison 

We will also explore the relationship between fireplacecnt and yearbuilt. As houses get newer, I believe they have less fireplaces.
```{r}
CraigsHousing.byYear = CraigsHousing %>% group_by(yearbuilt) %>%
  summarize(
    m.fireplacecnt = mean(fireplacecnt, na.rm=TRUE)
  )
CraigsHousing.byYear %>% ggplot(aes(x=yearbuilt, y=m.fireplacecnt)) + geom_line()
``` 

This data in a way confirms my idea that newer houses have less fireplaces, but they seem to come back in trends, such as around 200, and 2015 or so. This graph did definitely not look like I thought it would. 

### Linear regression
Now we will attempt a linear regression. In theory we should do variable selection and test p-values and such. Instead I am going to use yearbuilt, bathroomcnt, and regionidcity. bathroomcnt I believe will act as "how big is the house", while yearbuilt determines age. These are the factors I will use to try and explain taxvaluedollarcnt.

First we check the explanatory variables against the response.
```{r}
regressionData = HousingOriginal %>% select(taxvaluedollarcnt, yearbuilt, bathroomcnt)
res <- cor(regressionData)
round(res, 2)

regressionData %>% ggplot(aes(x=taxvaluedollarcnt, y=yearbuilt)) + geom_point()
regressionData %>% ggplot(aes(x=taxvaluedollarcnt, y=bathroomcnt)) + geom_point()
```

bathroomcnt is moderately visually linear. Yearbuilt looks around moderately as well. 

Next, as practice, we will try a single regression investigating whether using only yearbuilt helps to explain taxvaluedollarcnt.
```{r}
single.regression <- lm(taxvaluedollarcnt ~ yearbuilt, data = regressionData)

summary(single.regression)
```

Does this model help to explain taxvaluedollarcnt? 
Null hypothesis: b1 = 0
Alternative hypothesis: b1 != 0
f statistic: 2.456e+04
p value: <0.0001
Conclusion: we have overwhelimg evidence that using yearbuilt helps to explain the value of taxvaluedollarcnt in the sample.
Equation = -7.342e+06 + 3.962e+03*yearbuilt
adjusted R^2: 0.02399 This means that using yearbuilt we can explain 2.4% of the variance in taxvaluedollarcnt

With real estate being an interesting and dynamic market a single variable obviously wouldn't do very well. I was expecting it to perform better though. How about more variables?
```{r}
multiple.regression <- lm(taxvaluedollarcnt ~ yearbuilt + bathroomcnt, data = regressionData)

summary(multiple.regression)
```

Does this model help to explain taxvaluedollarcnt?
Null hypothesis: b1 = b2 = 0
Alternative hypothesis: atleast one of b1, b2 != 0
f statistic: 1.333e+05
p value: <0.0001
Conclusion: we have overwhelimg evidence that using yearbuilt and bathroomcnt help to explain the value of taxvaluedollarcnt in the sample.
Equation = 615,355.83 - 417.42 * yearbuilt + 280,238.92 * bathroomcnt
adjusted R^2: 0.2106 This means that using yearbuilt and bathroomcnt we can explain 21% of the variance in taxvaluedollarcnt

These results are suspicious. $280k seems like a lot per bathroom, and as houses get newer I would think they get more valuable but that does not appear to be the case. Our R^2 is quite low, but raised dramatically from the single variable.

### Ngoc's Questions:

Ngoc's variables: basementsqft, finishedsquarefeet15, landtaxvaluedollarcnt, typeconstructiontypeid, garagetotalsqft
<!-- numeric vs numeric -->  
1. What is the relationship between basementsqft, and garagetotalsqft 
<!-- numeric vs categorical -->
2. What is the relationship between finishedsquarefeet15 and airconditioningtypeid
<!-- categorical vs categorical -->
3. What is the relationship between  airconditioningtypeid and typeconstructiontypeid

4. My expectation for relationship?

5. is or how-far the expectation met

6. Choose and add another into the relationship.

7. What does this finding mean?